Automatic Matching of Bullet Lands

### 2016-04-28 revision/response.Rmd

15.  The authors would need to study (or at least discuss extensively) the ability of their method to capture the variability in bullet profiles over time (at interval of several dozen of bullets fired by the same weapon, when the weapon has been cleaned, or when the bullet was degraded by hitting a target). In addition, a dataset of 10 firearms (even consecutively manufactured) does not provide any information on the rarity of a particular bullet profile, and therefore does not address the second probability statement. Alternatively, the proposed metrics could be tested in a database searching situation, which has not been done.

***At the time of the submission of this manuscript, the 35 bullets from Hamby set 252 were the only ones accessible to the public. They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database. In the future more results will be available, but until then we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***

16. On a related subject, I believe that a random forest provides the probability that two bullets were fired by the same gun given the set of observations (Fig 19), not the probability of interest: the probability of two sets of observations given that the two bullets were fired by the same gun. The authors may want to revise some of the statements in their discussion and conclusion.

***HH: why should the probability of two sets of observations given that the two bullets were fired by the same gun be the probability of interest? I would agree that the random forest and any of its conclusions are restricted to how good the underlying 'database' is. The more representative the set of bullets, the better the predictions should be. It is not the purpose of this paper to find the probability of two sets of observations given that the two bullets were fired by the same gun. We made sure to remove any misleading statements that might suggest the latter probability. ***

@@

21. First, the term "land" is used in the title but not defined until p. 3.  I recommend changing the title, to ensure broad comprehenson; perhaps "Automatic Matching of Bullet Marks," or something of that kind.

***To clarify the title, we have modified the abstract to provide a brief definition of bullet lands.***

22. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.

***Grooves and striations are actually not synonyms in this case. The grooves are impressions induced by the rifling of the barrel, and properties associated with them are class characteristics not unique to individual gun barrels. We've taken care to ensure the distinction is clear between the striae of interest, which are extracted from the lands of the bullet (between grooves), and the grooves themselves.***

@@

***EH: Add typical width of groove + lands***

***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***

***HH: I am not sure about the width of a striation - the width depends on where it is measured (in both x and z direction). I'm sure there are definitions of that available. We might need some more literature.***

***EH: http://nvlpubs.nist.gov/nistpubs/jres/116/3/v116.n03.a01.pdf - "The number and width of striations in a land impression can vary depending on the size of the bullet and the characteristics of the barrel. For example, according to Biasotti’s study, the Smith and Wesson barrel with a land impression width of 2.4 mm has an average of 60 striations with total striation counts between 16 and 97 [13]. Observing the land impression images used in our study, 50 pixels (about 0.08 mm) is a reasonable choice for the average width of a structuring element."***

### 2016-05-03 imaging-paper.Rnw
Refactor some library code

library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)
@

@@

  densKM <- sm.density(bstats[,x][bstats$match], display="none", 
                       eval.points=xx, weights=NA, method="normal")
  densKNM <- sm.density(bstats[,x][!bstats$match], display="none", 
                       eval.points=xx, weights=NA, method="normal")
  dframe <- data.frame(var=x, x = xx, KM=densKM$estimate, KNM = densKNM$estimate)
  dframe <- rbind(data.frame(var=x, x = xr[1], KM=0, KNM=0),
        dframe, data.frame(var=x, x = xr[2], KM=0, KNM=0))
  dframe$order <- 1:nrow(dframe)
  dframe
}) 
library(reshape2)
dm <- melt(densities, measure.var=c("KM", "KNM"))
dm$var <- factor(dm$var, levels=features)
levels(dm$var)[3:4] <- c("#matches", "#non-matches")
levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
dm$varLabel <- dm$var
levels(dm$varLabel) <- c(
  "Consecutive Matching Striae (CMS)", 
  "Consecutive Non-Matching Striae (CNMS)",
  "#matches",
  "#non-matches",
  "Average difference (D)",
  "Sum of peaks (S)",
  "Cross-correlation function (ccf)")

@@

\begin{figure}[hbtp]
  \centering
<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
library(rpart)
library(rpart.plot)
library(RColorBrewer)
vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
names(bstats)[9:10] <- c("#matches", "#non-matches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
#cc <- read.csv("csvs/crosscuts-25.csv")
# cc$bullet <- gsub(".*//","",cc$path)
# cc$bullet <- gsub(".x3p", "", cc$bullet)
# bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
#                   bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]

@@

<<rforest, echo=FALSE, message=FALSE>>=
library(randomForest)
library(reshape2)

set.seed(20160105)
names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))

>>>>

library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)
library(reshape2)
library(randomForest)
@

@@

  densKM <- sm.density(bstats[,x][bstats$match], display="none", 
                       eval.points=xx, weights=NA, method="normal")
  densKNM <- sm.density(bstats[,x][!bstats$match], display="none", 
                       eval.points=xx, weights=NA, method="normal")
  dframe <- data.frame(var=x, x = xx, KM=densKM$estimate, KNM = densKNM$estimate)
  dframe <- rbind(data.frame(var=x, x = xr[1], KM=0, KNM=0),
        dframe, data.frame(var=x, x = xr[2], KM=0, KNM=0))
  dframe$order <- 1:nrow(dframe)
  dframe
})

dm <- melt(densities, measure.var=c("KM", "KNM"))
dm$var <- factor(dm$var, levels=features)
levels(dm$var)[3:4] <- c("#matches", "#non-matches")
levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
dm$varLabel <- dm$var
levels(dm$varLabel) <- c(
  "Consecutive Matching Striae (CMS)", 
  "Consecutive Non-Matching Striae (CNMS)",
  "#matches",
  "#non-matches",
  "Average difference (D)",
  "Sum of peaks (S)",
  "Cross-correlation function (ccf)")

@@

\begin{figure}[hbtp]
  \centering
<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
names(bstats)[9:10] <- c("#matches", "#non-matches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
#cc <- read.csv("csvs/crosscuts-25.csv")
# cc$bullet <- gsub(".*//","",cc$path)
# cc$bullet <- gsub(".x3p", "", cc$bullet)
# bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
#                   bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]

@@

<<rforest, echo=FALSE, message=FALSE>>=
set.seed(20160105)
names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
